### 2024-04-26 README.Rmd
readme

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.6077618.svg)](https://doi.org/10.5281/zenodo.6077618)

![Antigua fábrica de La Cerámica Gijonesa](ceramica.JPG)

# Anthropogenic plant communities of the Iberian Atlantic ecoregion

This repository stores all the information related to the manuscript [*Anthropogenic plant communities of the Iberian Atlantic ecoregion*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

## Contents

This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.

The following materials are available in the folders of this repository:

* `data` Data files including the vegetation plots.
* `doc` Files to create the [manuscript](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md) using `Rmarkdown`.
* `results` Output of the `R` scripts, including figures and model outputs.
* `src` Scripts in `R` language used to clean the raw data and perform the analyses of the manuscript.

## Abstract

* *Questions* Which is the diversity of anthropogenic plant communities at the ecoregional level in a transitional territory between temperate and Mediterranean Europe? How are these communities composed by native plants, archaeophytes and neophytes? What are the main environmental and disturbance drivers of the diversity in anthropogenic vegetation?

* *Location* Iberian Atlantic ecoregion, Iberian Peninsula, south-western Europe.

* *Methods* We compiled a vegetation database of 2,508 synanthropic plots and revised the regional checklist of anthropogenic alliances. We used modified TWINSPAN and semi-supervised classification to classify the plots into the revised alliances. We determined the proportion of natives, archaeophytes and neophytes. We described the alliances in terms of vegetation height, flowering phenology and ecological requirements for temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.

* *Results* We classified 2,081 vegetation plots into 28 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*). The plots included 1,162 plant taxa: 78% natives, 15% archaeophytes and 7% neophytes. The main ecological gradients behind this diversity went from dry-sunny to moist-shady habitats, followed by a gradient of low to high disturbance frequency and severity.

* *Conclusions* The human-made habitats of the Iberian Atlantic ecoregion host one third of the ecoregional species pool and one fifth of the Iberian flora. Wet anthropogenic habitats are most affected by neophyte invasions. Annual vegetation of arable lands is a refuge for archaeophytes. Mesic perennial ruderal vegetation is specially rich in native species and can be a biodiversity asset in urban and peri-urban settings.

## Citation

Please cite the repository, datasets and article as:

> Fernández-Pascual E, González-García, Ivesdal G, Lázaro-Lobo A, Jiménez-Alfaro B. [*Anthropogenic plant communities of the Iberian Atlantic ecoregion*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), *in preparation*.