### 2024-11-27 README.Rmd
updated README and GitHub manuscript 📖

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.11613571.svg)](https://zenodo.org/doi/10.5281/zenodo.11613571)

![](ceramica.JPG)

# Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula

This repository stores all the information related to the manuscript [*Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), including the datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

## Contents

This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.

The following materials are available in the folders of this repository:

* `data` Data files including the vegetation plots.
* `doc` Files to create the [manuscript](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md) using `Rmarkdown`.
* `results` Output of the `R` scripts, including figures and model outputs.
* `src` Scripts in `R` language used to clean the raw data and perform the analyses of the manuscript.

## Abstract

* *Questions* Can we reconcile regional and European classifications of anthropogenic plant communities at the biogeographical scale? How are these communities characterized by species origins, traits and ecological preferences?

* *Location* Atlantic territories in the NW Iberian Peninsula (a.k.a. Cantabrian Mixed Forests ecoregion); south-western Europe.

* *Methods* We classified 2,508 plots with the aim of being consistent with regional phytosociological expertise, while matching that expertise with current EuroVegChecklist alliances. We used modified TWINSPAN to revise the original phytosociological classification, followed by semi-supervised re-classification of the whole dataset. We determined the proportion of natives, archaeophytes and neophytes. We also described the alliances in terms of species traits (lifeforms, height and flowering phenology) and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency and severity).

* *Results* We assigned 2,086 vegetation plots to 25 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*). The plots included 1,149 species: 78% natives, 15% archaeophytes and 7% neophytes. Vegetation groups were organized along a principal axis of abiotic stress (dry-sunny to moist-shady habitats) and a secondary axis of disturbance.

* *Conclusions* In the Iberian Atlantic territories, anthropogenic habitats host one third of the regional plant species pool and one fifth of the Iberian flora. Mesic perennial ruderal vegetation is especially rich in native species and can be a biodiversity asset in urban landscapes. Our biogeographical-level synthesis can improve the management of anthropogenic plant communities and contribute towards a European-level synthesis of human-made vegetation.

## Citation

Please cite the repository, datasets and article as:

> Fernández-Pascual E, González-García, Ivesdal G, Lázaro-Lobo A, Jiménez-Alfaro B. [*Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), *in preparation*.